### 2017-07-20 Docs/makefiles/makefiles.tex

The main macros that define the build process are split across several
files.  The 4 main files are:
\begin{itemize}
\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakedefs.mak}:

@@

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:

@@

\item {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/GMakerules.mak}:

@@

  This is a \maestro-specific file that gathers all of the various
  modules that are used to build a typical \maestro\ application
  and integrates with the \boxlib\ build system.  Every \maestro\
  problem's {\tt GNUmakefile} will include this file.

@@

\maestro\ gets the location of the \boxlib\ library through the 
{\tt BOXLIB\_HOME} variable.  This should be set as an environment
variable in your shell start-up files (e.g.\ {\tt .bashrc} or {\tt
.cshrc}).

The \boxlib\ build system separates the compiler-specific information
from the machine-specific information---this allows for reuse of the
compiler information.  The only machine-specific parts of the build system
are for the MPI library locations, contained in {\tt GMakeMPI.mak}.
The compiler flags for the various compilers are listed in the
files in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.  The compiler
is set via the {\tt COMP} variable in the problem's {\tt GNUmakefile}.

@@

The \boxlib\ build system relies on the vpath functionality of {\tt
make}.  In a makefile, the {\tt vpath} variable holds search path used
to locate source files.  When a particular file is needed, the
directories listed in {\tt vpath} are searched until the file is
found.  The first instance of a file is used.  We exploit this feature
by always putting the build directory first in {\tt vpath} (this is
done in {\tt GMakerules.mak}).  This means that if a source file is
placed in the build directory, that copy will override any other
version in the source path.

@@

There is no need to explicitly define the dependencies between the
source files for Fortran modules.  The scripts in {\tt
BOXLIB\_HOME/Tools/F\_scripts/} are run at the start of the build
process and parse all the source files and make an explicit list of
the dependency pairs.  The execution of these scripts is triggered
by including makefiles of the form {\tt *.depends}.  On a fresh build these
will not exist.  When GNU make cannot find an included makefile it will
first attempt to build it using any relevant targets before issuing an 
error.  The targets for the {\tt *.depends} files contain the recipe for
executing the dependency scripts.  Once these makefiles are built by the
scripts GNU make will then read the dependencies for the current build
from them.  This process is defined in {\tt GMakerules.mak}.

@@

  This is the module that controls runtime parameters.  This is
  created by the script
  {\tt write\_probin.py} in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/}.  The
  makefile logic for creating it is in {\tt GMaestro.mak}.  At compile
  time, the problem, main {\tt MAESTRO/}, and any microphysics
  directories (set from the {\tt EOS\_DIR}, {\tt CONDUCTIVITY\_DIR}, and {\tt NETWORK\_DIR} parameters in the {\tt GNUmakefile}
  are searched for {\tt \_parameter} files.  These files
  are parsed and the {\tt probin.f90} file is output containing the 
  runtime parameters, their default values, and the logic for reading
  and interpreting the inputs file.

@@

  This is a module that contains basic information about the build
  environment (compiler used, build machine, build directory, compiler
  flags, etc.).  This is created by the script {\tt makebuildinfo.py}
  in {\tt \$\{BOXLIB\_HOME\}/Tools/F\_scripts/} from {\tt
  GMaestro.mak} by passing in a number of makefile variables.  This is
  rewritten everytime the code is compiled.  The primary use of this
  module is writing out the {\tt job\_info} file in the plotfiles.

@@

Several modules are included in all \maestro\ builds by default.
From \boxlib, we alway include:
\begin{itemize}
\item {\tt \$\{BOXLIB\_HOME\}/Src/F\_BaseLib}
\item {\tt \$\{BOXLIB\_HOME\}/Src/LinearSolvers/F\_MG}
\end{itemize}

@@

\subsection{\boxlib-only Tests}

An even more restrictive setting than {\tt UNIT\_TEST := t} is invoked
by setting {\tt BOXLIB\_ONLY := t}.  This is like the unit test flag,
but does not include {\tt MAESTRO/Source} in the {\tt vpath} search
path for sources.  So this is intended for cases where we don't want
to use any \maestro\ source files.  Typically, this is used in the
small unit tests that live under the various microphysics solvers.  If
a {\tt probin.f90} is built for these tests, it will not include all
the \maestro-specific parameters, but will include any parameters from
the various microphysics routines.

@@

\boxlib\ has several options that produce executables that can help
track down memory issues, uninitialized variables, NaNs, etc.

@@

  Setting {\tt TEST := t}\otherindex{{\tt GNUmakefile}}{TEST} will
  enable routines in \boxlib\ initialize multifabs and arrays 
  allowed via {\tt bl\_allocate} to signalliing NaNs.  This behavior
  is the same as {\tt NDEBUG :=}, but {\tt TEST := t} uses the 
  same compiler optimizations as a normal build.

  This can be useful with compiler flags that trap floating point
  exceptions (FPEs), but checks on floating point exceptions can also
  be enabled through runtime parameters passed to \boxlib's
  backtrace functionlity:
  \begin{itemize}
  \item \runparam{boxlib\_fpe\_invalid}: enabling FPE trapping for
    invalid operations (e.g. {\tt 0 * inf}, {\tt sqrt(-1)})

@@

  When exception trapping is enabled (either via \boxlib\ or the
  compiler), the code will abort, and the backtrace information will
  be output to a file {\tt Backtrace.N}, where {\tt N} is the
  processor number.  BoxLib will also initialize \multifab s with
  signaliing NaNs to help uncover any floating point issues.

@@

  Behind the scenes, \boxlib\ implements this capability via the
  Linux/Unix {\tt feenableexcept} function (this is in {\tt
    backtrace\_c.cpp} in \boxlib).

@@

Properties for different compilers are already defined in {\tt
  \$\{BOXLIB\_HOME\}/Tools/F\_mk/comps/}.  Each compiler is given its
own file.  The appropriate file is included into {\tt GMakedefs.mak}
by looking at the {\tt COMP} variable and the operating system.  These
compiler files define the compiler flags for both optimized and debug
compiling.  Additionally, the variable {\tt FCOMP\_VERSION} should be
defined there, based on the output from the compiler, to provide the
compiler version for output into the {\tt job\_info} file at runtime.

@@

When building with MPI, the build system needs to know about the location of the
MPI libraries.  If your local MPI has the {\tt mpif90} and {\tt mpicc} wrappers
installed and working, then \maestro\ can use these simply by setting the
environment variable:
\begin{verbatim}
BOXLIB_USE_MPI_WRAPPERS=1
\end{verbatim}
Otherwise, you will need to edit {\tt GMakeMPI.mak} and add a section
specific to your machine with the compiler and library location.  It
is best to simply copy an existing similar portion of the makefile and
adjust it to your system.  Most national supercomputing facilities are
already supported, and parallel builds on them should work out of the
box.

>>>>

The main macros that define the build process are split across several
files.  The 4 main files are:
\begin{itemize}
\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakedefs.mak}:

@@

\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakeMPI.mak}:

@@

\item {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/GMakerules.mak}:

@@

  This is a \maestro-specific file that gathers all of the various
  modules that are used to build a typical \maestro\ application
  and integrates with the \amrex\ build system.  Every \maestro\
  problem's {\tt GNUmakefile} will include this file.

@@

\maestro\ gets the location of the \amrex\ library through the 
{\tt AMREX\_HOME} variable.  This should be set as an environment
variable in your shell start-up files (e.g.\ {\tt .bashrc} or {\tt
.cshrc}).

The \amrex\ build system separates the compiler-specific information
from the machine-specific information---this allows for reuse of the
compiler information.  The only machine-specific parts of the build system
are for the MPI library locations, contained in {\tt GMakeMPI.mak}.
The compiler flags for the various compilers are listed in the
files in {\tt \$\{AMREX\_HOME\}/Tools/F\_mk/comps/}.  The compiler
is set via the {\tt COMP} variable in the problem's {\tt GNUmakefile}.

@@

The \amrex\ build system relies on the vpath functionality of {\tt
make}.  In a makefile, the {\tt vpath} variable holds search path used
to locate source files.  When a particular file is needed, the
directories listed in {\tt vpath} are searched until the file is
found.  The first instance of a file is used.  We exploit this feature
by always putting the build directory first in {\tt vpath} (this is
done in {\tt GMakerules.mak}).  This means that if a source file is
placed in the build directory, that copy will override any other
version in the source path.

@@

There is no need to explicitly define the dependencies between the
source files for Fortran modules.  The scripts in {\tt
AMREX\_HOME/Tools/F\_scripts/} are run at the start of the build
process and parse all the source files and make an explicit list of
the dependency pairs.  The execution of these scripts is triggered
by including makefiles of the form {\tt *.depends}.  On a fresh build these
will not exist.  When GNU make cannot find an included makefile it will
first attempt to build it using any relevant targets before issuing an 
error.  The targets for the {\tt *.depends} files contain the recipe for
executing the dependency scripts.  Once these makefiles are built by the
scripts GNU make will then read the dependencies for the current build
from them.  This process is defined in {\tt GMakerules.mak}.

@@

  This is the module that controls runtime parameters.  This is
  created by the script
  {\tt write\_probin.py} in {\tt \$\{AMREX\_HOME\}/Tools/F\_scripts/}.  The
  makefile logic for creating it is in {\tt GMaestro.mak}.  At compile
  time, the problem, main {\tt MAESTRO/}, and any microphysics
  directories (set from the {\tt EOS\_DIR}, {\tt CONDUCTIVITY\_DIR}, and {\tt NETWORK\_DIR} parameters in the {\tt GNUmakefile}
  are searched for {\tt \_parameter} files.  These files
  are parsed and the {\tt probin.f90} file is output containing the 
  runtime parameters, their default values, and the logic for reading
  and interpreting the inputs file.

@@

  This is a module that contains basic information about the build
  environment (compiler used, build machine, build directory, compiler
  flags, etc.).  This is created by the script {\tt makebuildinfo.py}
  in {\tt \$\{AMREX\_HOME\}/Tools/F\_scripts/} from {\tt
  GMaestro.mak} by passing in a number of makefile variables.  This is
  rewritten everytime the code is compiled.  The primary use of this
  module is writing out the {\tt job\_info} file in the plotfiles.

@@

Several modules are included in all \maestro\ builds by default.
From \amrex, we alway include:
\begin{itemize}
\item {\tt \$\{AMREX\_HOME\}/Src/F\_BaseLib}
\item {\tt \$\{AMREX\_HOME\}/Src/LinearSolvers/F\_MG}
\end{itemize}

@@

\subsection{\amrex-only Tests}

An even more restrictive setting than {\tt UNIT\_TEST := t} is invoked
by setting {\tt AMREX\_ONLY := t}.  This is like the unit test flag,
but does not include {\tt MAESTRO/Source} in the {\tt vpath} search
path for sources.  So this is intended for cases where we don't want
to use any \maestro\ source files.  Typically, this is used in the
small unit tests that live under the various microphysics solvers.  If
a {\tt probin.f90} is built for these tests, it will not include all
the \maestro-specific parameters, but will include any parameters from
the various microphysics routines.

@@

\amrex\ has several options that produce executables that can help
track down memory issues, uninitialized variables, NaNs, etc.

@@

  Setting {\tt TEST := t}\otherindex{{\tt GNUmakefile}}{TEST} will
  enable routines in \amrex\ initialize multifabs and arrays 
  allowed via {\tt bl\_allocate} to signalliing NaNs.  This behavior
  is the same as {\tt NDEBUG :=}, but {\tt TEST := t} uses the 
  same compiler optimizations as a normal build.

  This can be useful with compiler flags that trap floating point
  exceptions (FPEs), but checks on floating point exceptions can also
  be enabled through runtime parameters passed to \amrex's
  backtrace functionlity:
  \begin{itemize}
  \item \runparam{boxlib\_fpe\_invalid}: enabling FPE trapping for
    invalid operations (e.g. {\tt 0 * inf}, {\tt sqrt(-1)})

@@

  When exception trapping is enabled (either via \amrex\ or the
  compiler), the code will abort, and the backtrace information will
  be output to a file {\tt Backtrace.N}, where {\tt N} is the
  processor number.  \amrex\ will also initialize \multifab s with
  signaliing NaNs to help uncover any floating point issues.

@@

  Behind the scenes, \amrex\ implements this capability via the
  Linux/Unix {\tt feenableexcept} function (this is in {\tt
    backtrace\_c.cpp} in \amrex).

@@

Properties for different compilers are already defined in {\tt
  \$\{AMREX\_HOME\}/Tools/F\_mk/comps/}.  Each compiler is given its
own file.  The appropriate file is included into {\tt GMakedefs.mak}
by looking at the {\tt COMP} variable and the operating system.  These
compiler files define the compiler flags for both optimized and debug
compiling.  Additionally, the variable {\tt FCOMP\_VERSION} should be
defined there, based on the output from the compiler, to provide the
compiler version for output into the {\tt job\_info} file at runtime.

@@

When building with MPI, the build system needs to know about the
location of the MPI libraries.  If your local MPI has the {\tt mpif90}
and {\tt mpicc} wrappers installed and working, then \maestro\ will
attempt to use these.  Otherwise, you will need to edit {\tt
  GMakeMPI.mak} and add a section specific to your machine with the
compiler and library location.  It is best to simply copy an existing
similar portion of the makefile and adjust it to your system.  Most
national supercomputing facilities are already supported, and parallel
builds on them should work out of the box.
